---
title: "R Notebook"
output:
  html_document:
    df_print: paged
---

Q3:See the stopping rule below. Explain to your audience how you used computer simulation to estimate the average number of plays before stopping. The code below will need to be modified to calculate this quantity.

For Q3, we can see the second chunk code. In order to attain the average number of plays before stopping, we can know the number of the change of budget to find the number of plays. And by finding the average number of the change of budget to the average number of plays before stopping. By the two chunk code calculation, we can attain that the average number of plays before stopping is 203.396.

```{r}
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
walk_out_money <- rep(NA, 10000)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Walk out money distribution
hist(walk_out_money, breaks = 100)

# Estimated probability of walking out with extra cash
mean(walk_out_money > 200)

# Estimated earnings
mean(walk_out_money - 200)

```

```{r}
get_play <- function(x) length(x) #get_play(x) = length(x)
play_number <- rep(NA, 1000)
for(j in seq_along(play_number)){
  play_number[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_play
}
mean(play_number)


```

