---
title: "Analyzing roulette strategies"
author: "Qianyi Sun"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

#  Introduction

A roulette table composed of 38 (or 37) evenly sized pockets on a wheel. The pockets are colored red, black, or green. The pockets are also numbered. Roulette is a game of chance in which a pocket is randomly selected. Gamblers may wager on several aspects of the outcome. For example, one may place a wager that the randomly selected pocket will be red or odd numbered or will be a specific number.

For this game, all one needs to know is that there are 38 pockets of which 2 are green, 18 are red, and 18 are black. The payout for a bet on black (or red) is $1 for each $1 wagered. This means that if a gambler bets $1 on black and the randomly selected pocket is black, then the gambler will get the original $1 wager and an additional $1 as winnings.(quoted from https://github.com/thomasgstewart/data-science-5620/blob/master/deliverables/01-roulette.md)

# Solutions to questions

## Simulation process of the roulette wheel
Q1:You should explain how you used computer simulation to calculate the average earnings of a gambler that uses this strategy. As part of the explanation, provide a figure (or a series of figures) that show how the gamblers earnings (or losses) evolve over a series of wagers at the roulette wheel. (The x-axis will be the wager number (or play number), the y-axis will be earnings.) The code below provides all the functions you’ll need to calculate average earnings.

For Q1, first of all, to solve this question, we need to find the relative variables to create a figure to observe how gambler earnings evolve a series of wagers at the roulette wheel.Thus, we at least must know how to attain x-axis and y-axis. By reading question, we can see that x-axis is wager number and y-axis is earnings. Then for wager number, how can we get it? In order to get wager number, we should understand one_series() function. Within the one_series code, it set initial state like B number which is the starting budget and W number which is the budget threshold for successfully stoping. And it opens the vector to store budget over series of plays and create loop of plays. Thus in a word, one_series() is to set basic game rule and has contained the every change of budget's values. So when we get the every change of budget's values, we can Subtract it from the starting budget to get the change of gambler earning's value.
At this time, we can use plot() to draw a figure to solve Q1. The plot() is a drawing function. We can see the second chunk code to know a simple way to solve Q1.


```{r}
library(dplyr)
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
walk_out_money <- rep(NA, 10000)
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_last
}

# Walk out money distribution
hist(walk_out_money, breaks = 100,xlab = "Walk Out Money", main = paste("Histogram of", "Walk Out Money"))

# Estimated probability of walking out with extra cash
mean(walk_out_money > 200)

# Estimated earnings
mean(walk_out_money - 200)
```
```{r}
earnings <- one_series(B = 200, W = 300, L = 1000, M = 100) - 200


plot(earnings,type="l", xlab = "Wager Number", ylab = "Earnings")
```
From the graph, it shows how the gamblers earnings (or losses) evolve over a series of wagers at the roulette wheel.Every time th gambler wins, the gambler will earns only 1 dollar. But if the gambler losses all the time, the gambler will loss a lot money because the loss is exponentially increased and in the end the gambler will loss all the money he has.

## Changing parameter

Q2:Show your audience how changing a parameter of the simulation (see table below) does or does not have an impact on average earnings. A figure would be helpful.

For Q2, we can see that there are four parameters we can change. The four parameters are B, W, L and M. But these parameters are mutual restrains. For example, B is less than W. I set B=b. The value of b is from 201 to 300. we set walk_out_money as y-axis. walk_out_money means that when a gambler stops the game, how much money he or she has. And we set xb as x-axis. The values of xb is from 201 to 300. Then we use plot() to draw a figure. By analyzing the result from first chunk code, we can know that the starting budget(B)  and walk_out_money has positive correlation. At the same time, we can indirectly find that the probability of winning is higher when the starting budget(B) is closer to the budget threshold for successfully stoping(W). In the same way, we change the budget threshold for successfully stoping(W) in the second chunk code. From the figure, we can find that the budget threshold for successfully stoping and walk_out_money has negative correlation. And By changing the maximum number of plays(L) and the casino wager limit(M), we can find that L and M has no distinct correlation.

### changing B value
```{r}

#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the starting budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
library(dplyr)
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
store_outmoney <- rep(NA, 100)
for(i in 1:100){
b=i+200 # set a oringinal value of Starting budget

walk_out_money <- rep(NA, 1000)#when walking out, person has the money ///if we want to run a plot quickly, we should set with rep()'s time is smaller interger
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = b, W = 300, L = 1000, M = 100) %>% get_last#get_last is function(x) x[length(x)] 's value of last one

}
store_outmoney[i] <-mean(walk_out_money)#store_outmoney contains all of mean of walk_out_money




# Estimated earnings
y=mean(walk_out_money - 200)
}
xb<- 201:300 # original budget is from 201 to 300.
plot(x=xb,y=store_outmoney,type="l",xlab="Starting budget",ylab="Average earnings") #draw a line chart
```
#### Figure analysis for changing B value
The graph shows that Starting budget and average earnings has positive relation.

### changing W value


```{r}
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the starting budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
library(dplyr)
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
store_outmoney <- rep(NA, 100)
for(i in 1:100){
w=i+200 # set a oringinal value of Starting budget

walk_out_money <- rep(NA, 1000)#when walking out, person has the money ///if we want to run a plot quickly, we should set with rep()'s time is smaller interger
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = w, L = 1000, M = 100) %>% get_last#get_last is function(x) x[length(x)] 's value of last one

}
store_outmoney[i] <-mean(walk_out_money)#store_outmoney contains all of mean of walk_out_money





# Estimated earnings
y=mean(walk_out_money - 200)
}
xw<- 201:300 # the budget threshold for successfully stoping is from 201 to 300.
plot(x=xw,y=store_outmoney,type="l",xlab="The budget threshold for successfully stoping", ylab="Average earnings")#draw a line chart
```
#### Figure analysis for changing W value
Then the plot will show the budget threshold for successfully stoping and average earnings has negative relation. It means that when the budget threshold for successfully stoping increases, average earnings have a tendency to decrease.

### changing L value

```{r}
#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the starting budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
library(dplyr)
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
store_outmoney <- rep(NA, 100)
for(i in 1:100){
l=i+900 # set a oringinal value of Starting budget

walk_out_money <- rep(NA, 1000)#when walking out, person has the money ///if we want to run a plot quickly, we should set with rep()'s time is smaller interger
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = l, M = 100) %>% get_last#get_last is function(x) x[length(x)] 's value of last one

}
store_outmoney[i] <-mean(walk_out_money)#store_outmoney contains all of mean of walk_out_money





# Estimated earnings
y=mean(walk_out_money - 200)
}
xl<- 901:1000 # the maximum number of plays is from 901 to 1000.
plot(x=xl,y=store_outmoney,type="l",xlab="The maximum number of plays", ylab = "Average earnings")#draw a line chart
```
#### Figure analysis for changing L value
The maximum number of plays and average earnings do not have a clear relation.


### changing M value

```{r}

#' A single play of the Martingale strategy
#'
#' Takes a state list, spins the roulette wheel, returns the state list with updated values (for example, budget, plays, etc)
#' @param state A list with the following entries: 
#'   B              number, the starting budget
#'   W              number, the budget threshold for successfully stoping
#'   L              number, the maximum number of plays 
#'   M              number, the casino wager limit
#'   plays          integer, the number of plays executed
#'   previous_wager number, the wager in the previous play (0 at first play)
#'   previous_win   TRUE/FALSE, indicator if the previous play was a win (TRUE at first play)
#' @return The updated state list
library(dplyr)
one_play <- function(state){
  
    # Wager
    proposed_wager <- ifelse(state$previous_win, 1, 2*state$previous_wager)
    wager <- min(proposed_wager, state$M, state$B)
    
    # Spin of the wheel
    red <- rbinom(1,1,18/38)
    
    # Update state
    state$plays <- state$plays + 1
    state$previous_wager <- wager
    if(red){
      # WIN
      state$B <- state$B + wager
      state$previous_win <- TRUE
    }else{
      # LOSE
      state$B <- state$B - wager
      state$previous_win <- FALSE
    }
  state
}


#' Stopping rule
#'
#' Takes the state list and determines if the gambler has to stop
#' @param state A list.  See one_play
#' @return TRUE/FALSE
stop_play <- function(state){
  if(state$B <= 0) return(TRUE)
  if(state$plays >= state$L) return(TRUE)
  if(state$B >= state$W) return(TRUE)
  FALSE
}


#' Play roulette to either bankruptcy, success, or play limits
#'
#' @param B number, the starting budget
#' @param W number, the budget threshold for successfully stoping
#' @param L number, the maximum number of plays 
#' @param M number, the casino wager limit
#' @return A vector of budget values calculated after each play.
one_series <- function(
    B = 200
  , W = 300
  , L = 1000
  , M = 100
){

  # initial state
  state <- list(
    B = B
  , W = W
  , L = L
  , M = M
  , plays = 0
  , previous_wager = 0
  , previous_win = TRUE
  )
  
  # vector to store budget over series of plays
  budget <- rep(NA, L)
  
  # For loop of plays
  for(i in 1:L){
    new_state <- state %>% one_play
    budget[i] <- new_state$B
    if(new_state %>% stop_play){
      return(budget[1:i])
    }
    state <- new_state
  }
  budget    
}

# helper function
get_last <- function(x) x[length(x)] 


# Simulation
store_outmoney <- rep(NA, 100)
for(i in 1:100){
m=i+50 # set a oringinal value of Starting budget

walk_out_money <- rep(NA, 1000)#when walking out, person has the money ///if we want to run a plot quickly, we should set with rep()'s time is smaller interger
for(j in seq_along(walk_out_money)){
  walk_out_money[j] <- one_series(B = 200, W = 300, L = 1000, M = m) %>% get_last#get_last is function(x) x[length(x)] 's value of last one

}
store_outmoney[i] <-mean(walk_out_money)#store_outmoney contains all of mean of walk_out_money





# Estimated earnings
y=mean(walk_out_money - 200)
}
xm<- 51:150 # number, the casino wager limit is from 51 to 150.
plot(x=xm,y=store_outmoney,type="l",xlab="The casino wager limit",ylab="Average earnings")#draw a line chart
```
#### Figure analysis for changing M value
The casino wager limit and average earnings do not have a clear relation.

## Estimating the average number of plays before stopping

Q3:See the stopping rule below. Explain to your audience how you used computer simulation to estimate the average number of plays before stopping. The code below will need to be modified to calculate this quantity.

For Q3, we can see the second chunk code. In order to attain the average number of plays before stopping, we can know the number of the change of budget to find the number of plays. And by finding the average number of the change of budget to the average number of plays before stopping. By the two chunk code calculation, we can attain that the average number of plays before stopping is 203.396.

```{r}
get_play <- function(x) length(x) #get_play(x) = length(x)
play_number <- rep(NA, 1000)
for(j in seq_along(play_number)){
  play_number[j] <- one_series(B = 200, W = 300, L = 1000, M = 100) %>% get_play
}
mean(play_number)
```

## Limitations,identify simplifications and uncertainty

Q4:Be sure to explain the limitations of the simulation; identify simplifications or other sources of uncertainty.

Limitations of the simulation:(1)When the number of gamblers rises above 10,000 times, the computer is running too long. (2)	When Starting budget(B) is closer to Winnings threshold for stopping(W), the gamblers is easier to win money at the end of game. 
Identifing simplifications:(1) Gamblers must bet until they meet stopping conditions and cannot quit the game on their own initiative. At the same time, casinos can't ignore the stopping rules and interrupt the game.
Uncertainty: we can not ignore the subjective factors of gambler which might affect the process of game.

