---
title: "05-quantile-precision"
date: "09/23/2019"
author: "Qianyi Sun"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: inline

---


# Introduction

As we all know, the median has an important meaning in doing some data analysis project. This is because the middle value of the data distribution can be used to indirectly represent data overall situation. But how to evaluate the median is very difficult.Thus now we begin to solve this problem.


For the distribution I need to calculate, we have already known that the random samples of size is N and the empirical sampling distribution should be at least 5000 draws. Then I need to Calculate the length of the middle 95% of the sampling distribution by calculating the difference between 2.5% and 97.5% percentiles of the sampling distribution.When I successfully calculate  the length of the middle 95% of the sampling distribution, I still need to calculate the density (relative likelihood) corresponding to the pt**h quantiles.

# Problem solutions
## Standard normal distribution
### Calculate the length of the middle 95% of the sampling distribution
```{r}
N=200
draws=5000
mid.len1<-rep(NA,91)  
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=91)
for(j in 1:91){
for(i in 1:draws){
  dist<-rnorm(N,mean=0,sd=1)
  pthquan[i]<-quantile(dist,p[j])
  
}
mid.len1[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}
plot(p,mid.len1,main="the length of the middle 95% of the sampling dist ",xlab="pth quantile",ylab = "Length")
```

When pth quantile is approaching the range between 0.4 and 0.6, the length of the middle 95% has nearly smallest value. Thus the range between 40th quantile and 60th quantile has more precision. However, when pth quantile is approaching 0th or 100th, the length of the middle 95% has higher value. Thus the 0th or 100th has less precision.
### Calculate the density corresponding to the pt**h quantiles 
```{r}
x_value<-rep(NA,91)
x_density<-rep(NA,91)
for (i in 1:91){
  d=i*0.01+0.04
  x_value[i]<-qnorm(d)
}
for (i in 1:91){
x_density[i]<-dnorm(x_value[i])
}
plot(x_density,mid.len1,xlab="pth quantile",ylab="Length",main="the length of the middle 95% of the sampling dist by density " )
```

From the graph, we can know that with the increase of pth quantile, the length is smaller. Thus the density is approaching to 0.4, the length has more precision. However, when the density is closer to 0.1, the length is less precisson.

## The exponential distribution with rate = 1

### Calculate the length of the middle 95% of the sampling distribution
```{r}
N=200
draws=5000
mid.len2<-rep(NA,91)   
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=91)
for(j in 1:91){
for(i in 1:draws){
  dist<-rexp(N, rate=1) 
  pthquan[i]<-quantile(dist,p[j])
  
}
mid.len2[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}
plot(p,mid.len2,main="the length of the middle 95% of the sampling dist ",xlab="pth quantile",ylab = "Length")
```

When pth quantile is approaching the 100th quantile, the length of the middle 95% has nearly biggest value. Thus the 100th has less precision. However, when pth quantile is approaching 0th, the length of the middle 95% has higher value. Thus the 0th has more precision.
### Calculate the density corresponding to the pt**h quantiles 

```{r}
x_value_2<-rep(NA,91)
x_density_2<-rep(NA,91)
for (i in 1:91){
  d=i*0.01+0.04
  x_value_2[i]<-qexp(d)
}
for (i in 1:91){
x_density_2[i]<-dexp(x_value_2[i])
}
plot(x_density_2,mid.len2,xlab="pth quantile",ylab="Length",main="the length of the middle 95% of the sampling dist by density ")
```

From the graph, we can know that with the decrease of pth quantile, the length is smaller. Thus the pth quantile is approaching to 1, the length has more precision. However, when the density is closer to 0.1, the length is less precisson.

## The mixture distribution

### Calculate the length of the middle 95% of the sampling distribution

```{r}

rf3 <- function(N){
  G <- sample(0:2, N, replace = TRUE, prob = c(5,3,2))
  (G==0)*rnorm(N) + (G==1)*rnorm(N,4) + (G==2)*rnorm(N,-4,2)
}

pf3 <- function(x){
  .5*pnorm(x) + .3*pnorm(x,4) + .2*pnorm(x,-4,2)
}

df3 <- function(x){
  .5*dnorm(x) + .3*dnorm(x,4) + .2*dnorm(x,-4,2)
}
draws=5000
mid.len3<-rep(NA,91)   #why it has a red mark
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=91)
for(j in 1:91){
for(i in 1:draws){
  dist<-rf3(200)
  pthquan[i]<-quantile(dist,p[j])
  
}
mid.len3[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}
plot(p,mid.len3,main="the length of the middle 95% of the sampling dist ",xlab="pth quantile",ylab = "Length")
```

When pth quantile is approaching the 40th quantile or 95th quantile, the length of the middle 95% has nearly smallest value. Thus the 40th quantile or 95th quantile has more precision. However, when pth quantile is approaching 10th, the length of the middle 95% has higher value. Thus the 0th has less precision.

### Calculate the density corresponding to the pt**h quantiles 
```{r}
p<-seq(0.05,0.95,length=91)
N=200
x_density<-rep(NA,91)
df3_3<-rep(NA,91)

for( i in 1:91){
  f1<-function(x){.5*pnorm(x) + .3*pnorm(x,4) + .2*pnorm(x,-4,2)-p[i]}
  x_density[i]<-uniroot(f1, lower = -200, upper = 200) 
} 

df3 <- function(x){
  .5*dnorm(x) + .3*dnorm(x,4) + .2*dnorm(x,-4,2)
}
x_density
for (i in seq_along(x_density)){
df3_3[i] <-df3(x_density[[i]])
}
df3_3 
plot(df3_3,mid.len3,xlab="pth quantile",ylab="Length",main="the length of the middle 95% of the sampling dist by density ")
```

From the graph, we can know that with the decrease of pth quantile, the length is smaller. Thus the density is approaching to 0.2, the length has more precision. However, when the density is closer to 0.05, the length is less precisson.

## The mixture distribution 4
### Calculate the length of the middle 95% of the sampling distribution

```{r}
rf4 <- function(N){
  G <- sample(0:1, N, replace = TRUE)
  (G==0)*rbeta(N,5,1) + (G==1)*rbeta(N,1,5)
}
draws=5000
mid.len4<-rep(NA,91)
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=91)
for(j in 1:91){
for(i in 1:draws){
  dist<-rf4(200)
  pthquan[i]<-quantile(dist,p[j])
  
}
mid.len4[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}
plot(p,mid.len4,main="the length of the middle 95% of the sampling dist ",xlab="pth quantile",ylab = "Length")
```

When pth quantile is approaching the 40th quantile or 95th quantile, the length of the middle 95% has nearly smallest value. Thus the 40th quantile or 95th quantile has more precision. However, when pth quantile is approaching 10th, the length of the middle 95% has higher value. Thus the 0th has less precision.

### Calculate the density corresponding to the pt**h quantiles 

```{r}
pf4<-function(x){
  0.5*pbeta(x,5,1)+0.5*pbeta(x,1,5)
}
p<-seq(0.05,0.95,length=91)
N=200
x_density_4<-rep(NA,91)
df4_1<-rep(NA,91)

for( i in 1:91){
  f4<-function(x){0.5*pbeta(x,5,1)+0.5*pbeta(x,1,5)-p[i]}
  x_density_4[i]<-uniroot(f4, lower = -200, upper = 200)
}  

df4 <- function(x){
  0.5*dbeta(x,5,1)+0.5*dbeta(x,1,5)
}
for (i in seq_along(x_density_4)){
df4_1[i] <-df4(x_density_4[[i]])
}
df4_1
plot(df4_1,mid.len4,,xlab="pth quantile",ylab="Length",main="the length of the middle 95% of the sampling dist by density ")
```

From the graph, we can know that with the decrease of pth quantile, the length is smaller. Thus the density is approaching to 2, the length has more precision. However, when the density is closer to 0.25, the length is less precisson.



# When does the median have the tightest sampling distribution?
```{r}
mid.len1[41]
mid.len2[41]
mid.len3[41]
mid.len4[41]
```

By see the four values, we can know the median of exponential distribution with rate = 1 has smallest value and thus has the tightest sampling distribution.

# Adding additional lines for selected distribution

In this point, I choose the standard normal distribution as the distribution to create a figure similar to the first one above, adding additional lines for N=400,800,1600.
```{r}
N=200
draws=5000
mid.len1<-rep(NA,19)   #why it has a red mark
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=19)
for(j in 1:19){
for(i in 1:draws){
  dist<-rnorm(N,mean=0,sd=1)
  pthquan[i]<-quantile(dist,p[j])
}
mid.len1[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}

N=400
mid.len1_2<-rep(NA,19)   #why it has a red mark
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=19)
for(j in 1:19){
for(i in 1:draws){
  dist<-rnorm(N,mean=0,sd=1)
  pthquan[i]<-quantile(dist,p[j])
}
mid.len1_2[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}

N=800
mid.len1_3<-rep(NA,19)   #why it has a red mark
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=19)
for(j in 1:19){
for(i in 1:draws){
  dist<-rnorm(N,mean=0,sd=1)
  pthquan[i]<-quantile(dist,p[j])
}
mid.len1_3[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}

N=1600
mid.len1_4<-rep(NA,19)   #why it has a red mark
pthquan<-rep(NA,draws)
p<-seq(0.05,0.95,length=19)
for(j in 1:19){
for(i in 1:draws){
  dist<-rnorm(N,mean=0,sd=1)
  pthquan[i]<-quantile(dist,p[j])
  
}
mid.len1_4[j]<-quantile(pthquan,0.975)-quantile(pthquan,0.025)
}

plot(p,mid.len1,type = "l",ylim=c(0,0.6),main="the length of the middle 95% of the sampling dist ",xlab="pth quantile",ylab = "Length")
lines(p,mid.len1_2,type="l",col="red") 
lines(p,mid.len1_3,type="l",col="green")
lines(p,mid.len1_4,type="l",col="blue")

text(0.05,0.6,"N=200",2,col="black")
text(0.05,0.4,"N=400",2,col="red")
text(0.01,0.3,"N=800",2,col="green")
text(0.01,0.2,"N=1600",2,col="blue")
```


## Analysing the graph above
By analyzing the graph above, we can know that when N is larger, the length of the middle 95% of the sampling distribution is less and has the more tight sampling distribution. It also means that the N is larger, the precision is more.





