---
title: "Analyzing world series home field"
author: "Qianyi Sun"
output:
  html_notebook:
    code_folding: hide
    highlight: tango
    number_sections: yes
    theme: united
    toc: yes
    toc_depth: 3
    toc_float: yes
  html_document:
    df_print: paged
    toc: yes
    toc_depth: '3'
---

#  Introduction
The home field advantage is when one team has a game in their own home stadium, they can have a higher probability of winning the game.In this blog, we analyze two team whose names are Braves and Yankees. Braves's home stadium is Atlanta and Yankees's home stadium is New York City(NYC = New York City, ATL = Atlanta). 


```{r}
#original
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
hfi <- c(0,0,1,1,1,0,0) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

# P_B
pb <- 0.55
advantage_multiplier <- 1.1 # Set = 1 for no advantage
pbh <- 0.55*advantage_multiplier
pba <- 1 - (1 - 0.55)*advantage_multiplier

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]
```


```{r}
#Q1_with_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
hfi <-  c(0,0,1,1,1,0,0) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

# P_B
pb <- 0.55
advantage_multiplier <- 1.1 # Set = 1 for no advantage
pbh <- 0.55*advantage_multiplier # have advantage
pba <- 1 - (1 - 0.55)*advantage_multiplier #no

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]
```
```{r}
#Q1_without_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")



# P_B
p_win<- 0.55
#advantage_multiplier <- 1.1 # Set = 1 for no advantage
#pbh <- 0.55*advantage_multiplier
#pba <- 1 - (1 - 0.55)*advantage_multiplier

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    #p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]

```
Answer: The difference in probabilites (with vs without home field advantage) does depend on the advantage factor. The conclusion comes from the answer of question 1. Through comparing with home field advantage result with without home field advantage result, the first result's probability of overall World Series outcomes that Braves can win is 0.604221. Otherwise, the second's probability of overall World Series outcomes that Braves can win is 0.6082878. The two values are different.


```{r}
#Q2_with_homeadvantage

require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
hfi <-  c(0,0,1,1,1,0,0) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

# P_B
pb <- mean(rbinom(10000,1000,0.55)/1000)
advantage_multiplier <- 1.1 # Set = 1 for no advantage
pbh <- mean(rbinom(10000,1000,0.55)/1000)*advantage_multiplier
pba <- 1 - (1 - mean(rbinom(10000,1000,0.55)/1000))*advantage_multiplier

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]

```





```{r}
#Q2_without_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator

# P_B
p_win<- mean(rbinom(10000,1000,0.55)/1000)
#advantage_multiplier <- 1.1 # Set = 1 for no advantage
#pbh <- 0.55*advantage_multiplier
#pba <- 1 - (1 - 0.55)*advantage_multiplier

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    #p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
apo[, sum(p), overall_outcome]

```




```{r}
#Q3
p_win3<-0.55
absolute_error_assign4<-abs(p_win3-mean(rbinom(10000,1000,0.55)/1000))
absolute_error_assign4

relative_error_assign4<-abs(p_win3-mean(rbinom(10000,1000,0.55)/1000))/p_win3
relative_error_assign4

```

```{r}

#Q4_with_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
hfi <-  c(0,0,1,1,1,0,0) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

# P_B
pb_win<-rep(NA,100)
pb_vector<-rep(NA,100)
for (l in 1:100){
  pb_win[l]<-0.01*l

advantage_multiplier <- 1.1 # Set = 1 for no advantage
pbh <- pb_win[l]*advantage_multiplier
pba <- 1 - (1 - pb_win[l])*advantage_multiplier

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win4 <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win4
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win4
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
d<-apo[, sum(p), overall_outcome]
pb_vector[l]<-d[[1,2]]

}


#Q4_without_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
#hfi <- c(1,1,0,0,0,1,1) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

pb_win_without<-rep(NA,100)
pb_vector_without<-rep(NA,100)
for (l in 1:100){
  pb_win_without[l]<-0.01*l
  
# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    #p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ pb_win_without[l]
      , apo[i,j,with=FALSE] == "L" ~ 1 - pb_win_without[l]
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
d_without<-apo[, sum(p), overall_outcome]
pb_vector_without[l]<-d_without[[1,2]]

}



```



```{r}
#with_home_advantage
v<-seq(0.01,1,by=0.01)
plot(x=v,y=pb_vector,type="l",xlab="PB",ylab="Probability of overalloutcomes that Braves wins",main="With home advantage")

```

```{r}
#without_home_advantage
v<-seq(0.01,1,by=0.01)
plot(x=v,y=pb_vector_without,type="l",xlab="PB",ylab="Probability of overalloutcomes that Braves wins",main="Without home advantage")
```


```{r}
pb_difference<-rep(NA,100)
for(i in 1:100){
pb_difference[i]<-pb_vector[i]-pb_vector_without[i]
}
v<-seq(0.01,1,by=0.01)
plot(x=v,y=pb_difference,type="l")

```

## Question 5

```{r}
#Q5_with_homeadvantage
require(dplyr)
require(data.table)
# Get all possible outcomes
apo <- fread("all-possible-world-series-outcomes.csv")

# Home field indicator
hfi <-  c(0,0,1,1,1,0,0) #{NYC, NYC, ATL, ATL, ATL, NYC, NYC}

# P_B
advantage_multiplier<-rep(NA,100)
pb_vector_5<-rep(NA,100)
for (l in 1:100){
 advantage_multiplier_C[l]<-1+l*0.01


pbh <- 0.55*advantage_multiplier_C[l]
pba <- 1 - (1 - 0.55)*advantage_multiplier_C[l]

# Calculate the probability of each possible outcome
apo[, p := NA_real_] # Initialize new column in apo to store prob
for(i in 1:nrow(apo)){
  prob_game <- rep(1, 7)
  for(j in 1:7){
    p_win <- ifelse(hfi[j], pbh, pba)
    prob_game[j] <- case_when(
        apo[i,j,with=FALSE] == "W" ~ p_win
      , apo[i,j,with=FALSE] == "L" ~ 1 - p_win
      , TRUE ~ 1
    )
  }
  apo[i, p := prod(prob_game)] # Data.table syntax
}

# Sanity check: does sum(p) == 1?
apo[, sum(p)] # This is data.table notation

# Probability of overall World Series outcomes
d<-apo[, sum(p), overall_outcome]
pb_vector_5[l]<-d[[1,2]]

}


```



```{r}
#Q5
pb_difference_5<-rep(NA,100)
pb_vector_without<-0.6082878
for(i in 1:100){
pb_difference_5[i]<-pb_vector_5[i]-pb_vector_without
}
v<-seq(0.01,1,by=0.01)
plot(x=v,y=pb_difference_5,type="l")
```





