---
title: "09-simulation-study"
author: "Qianyi Sun"
output: 
  html_document:
    theme: united
    #highlight: zenburn
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    #number_sections: true
    fig_caption: true
---

# Introduction
In this assignment, you will perform a 2×4×2 factorial simulation study to compare the coverage probability of various methods of calculating 90% confidence intervals. The three factors in the experiment are

True, underlying distribution
standard normal
gamma(shape = 1.4, scale = 3)
Model
method of moments with normal
method of moments with gamma
kernel density estimation
bootstrap
Parameter of interest
sample min (1st order statistic)
median
Other settings in the experiment that will not change are:

Sample size, N=201
Outside the loop estimation

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
```

```{r}
library(dplyr)
require(foreach)
require(doParallel)
library(tidyverse)
```

```{r}
data <- rnorm(201)

generate_data <- function(N, dist, sh, sc) {
  if (dist == "norm") {
    rnorm(N)
  } else if (dist == "gamma") {
    rgamma(N, shape = sh, scale = sc)
  }
}

estimate.ci <- function(data, mod, par.int, R = 10, smoo = 0.3) { #Parameter of interest
  N = length(data)
  sum.measure <- get(par.int)
  
  if (mod == "MMnorm") {
    
    mm.mean <- mean(data)
    mm.sd <- sd(data)
    
    sim.data <- array(rnorm(length(data)*R), dim = c(N, R)) #dim =c(N,R) produce N row
    samp.dist <- apply(sim.data, 2, FUN = sum.measure) # MARGIN: row is 1 row calucation
    return(quantile(samp.dist, c(0.05, 0.95)))
    
  } else if (mod == "MMgamma") {
    
    mm.shape <- mean(data)^2 / var(data)
    mm.scale <- var(data) / mean(data)
    
    sim.data <- array(rgamma(length(data)*R, shape = mm.shape, scale = mm.scale), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95),na.rm = TRUE)) #here I add na.rm=TRUE
    
  } else if (mod == "KDE") {
    
    ecdfstar <- function(t, data, smooth = smoo) {
      outer(t, data, function(a,b){pnorm(a, b, smooth)}) %>% rowMeans
    }
    
    tbl <- data.frame(
      x = seq(min(data)-sd(data), max(data)+sd(data), by = 0.01)
    ) 
    
    tbl$p <- ecdfstar(tbl$x, data, smoo) 
    tbl <- tbl[!duplicated(tbl$p), ]
    
    qkde <- function(ps, tbl){
      rows <- cut(ps, tbl$p, labels = FALSE) 
      tbl[rows, "x"]
    }
    
    U <- runif(N*R)
    sim.data <- array(qkde(U, tbl), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95),na.rm = TRUE))
    
  } else if (mod == "Boot") {
    
    samp.dist <- NA
    for (i in 1:R) {
      sim.data <- sample(data, N, replace = TRUE)
      if(par.int == "median") {
        samp.dist <- median(sim.data)
      } else if (par.int == "min") {
        samp.dist <- min(sim.data)
      }
    }
    return(quantile(samp.dist, c(0.05, 0.95),na.rm = TRUE))
    
    
    sim.data <- array(sample(data, N*R, replace = TRUE), dim = c(N, R))
    samp.dist <- apply(sim.data, 2, FUN = sum.measure)
    return(quantile(samp.dist, c(0.05, 0.95),na.rm = TRUE))
  }
}

capture_par <- function(ci, true.par) {
  1 * (ci[1] < true.par & true.par < ci[2])
}
```

# Run function and attain results

```{r}
N <- 201
shape.set <- 1.4
scale.set <- 3

true.norm.med <- qnorm(0.5)
true.norm.min <- mean(apply(array(rnorm(N*10000), dim = c(N, 10000)), 2, min))
true.gamma.med <- qgamma(0.5, shape = shape.set, scale = scale.set)
true.gamma.min <- mean(apply(array(rgamma(N*10000, shape = shape.set, scale = scale.set), dim = c(N, 10000)), 2, min))



simsettings <- expand.grid(dist = c("norm", "gamma"), model = c("MMnorm", "MMgamma", "KDE", "Boot"), par.int = c("median", "min"), cov.prob = NA, stringsAsFactors = FALSE, KEEP.OUT.ATTRS = FALSE) 

for (k in 1: nrow(simsettings)) {
  dist1 <- simsettings[k, 1]
  model1 <- simsettings[k, 2]
  par.int1 <- simsettings[k, 3]
  
  if (dist1 == "norm"& par.int1 == "median") {
    true.par1 = true.norm.med
  } else if (dist1 == "norm"&par.int1=="min") {
    true.par1 =true.norm.min
  }else if (dist1 =="gamma" & par.int1=="median"){
    true.part1=true.gamma.med
  }else if (dist1=="gamma" & par.int1=="min"){
    true.part1=true.gamma.min
  }
  
cover <- NA
for (sims in 1:10) {
  cover[sims] <- generate_data (N, dist1,sh=shape.set,sc=scale.set)%>% estimate.ci(mod = model1, par.int = par.int1) %>% capture_par(true.par = true.par1)
}
simsettings[k,4] <- mean(cover)
} 
simsettings
```

# Conclusion
By calculating, we can know that not all methods can work. So we can see that some answer is not correct and their coverage probability is very low.

