---
title: "09-method-comparison"
output:
  html_document:
    code_folding: hide
    theme: cosmo
    toc: yes
    toc_depth: 5
    toc_float: true
editor_options: 
  chunk_output_type: inline
---

# Introduction
In this assignment, you will perform a 2×4×2 factorial simulation study to compare the coverage probability of various<br> methods of calculating 90% confidence intervals. The three factors in the experiment are:

1.True, underlying distribution:<br>
standard normal<br>
gamma(shape = 1.4, scale = 3)<br>
2.Model:<br>
method of moments with normal<br>
method of moments with gamma<br>
kernel density estimation<br>
bootstrap<br>
3.Parameter of interest<br>
sample min (1st order statistic)<br>
median<br>

Other settings in the experiment that will not change are:<br>
Sample size, N=201<br>
Outside the loop estimation<br>

```{r setup, include=FALSE}
knitr::opts_chunk$set(warning = FALSE, message = FALSE, cache = TRUE)
```

```{r}
library(dplyr)
require(foreach)
require(doParallel)
library(tidyverse)
```

```{r}
generate_data<-function(N,dist,sh=1.4,sc=3){
  if (dist=="norm"){
    rnorm(N)+4
  }else if(dist=='gamma'){
    rgamma(N,shape=sh,scale=sc)
  }
}



estimate.ci<-function(data,mod,par.int,R=100,smoo=0.3){
  N<-length(data)
  sum.measure<-get(par.int)
  if (mod=="MMnorm"){ 
    mm.mean<-mean(data)
    mm.sd<-sd(data)
    
    
    sim.data<-array(rnorm(N*R,mm.mean,mm.sd),dim=c(N,R))  
    
  }
  else  if(mod=="MMgamma"){
    mm.shape<-mean(data)^2/var(data)
    mm.scale<-var(data)/mean(data)
    
    sim.data<-array(rgamma(N*R,shape=mm.shape,scale=mm.scale),dim=c(N,R)) 
    
    
  }
  else if (mod=="KDE"){
    
    ecdfstar <- function(t, data, smooth=smoo){
      outer(t, data, function(a,b){ pnorm(a, b, smooth)}) %>% rowMeans
    }
    
    
    tbl <- data.frame(
      x = seq(min(data)-sd(data),max(data)+sd(data),by = 0.01)
    )
    
    
    tbl$p <- ecdfstar(tbl$x, data) 
    
    tbl <- tbl[!duplicated(tbl$p),]
    
    qkde <- function(ps, tbl){
      rows <- cut(ps, tbl$p, labels = FALSE)
      tbl[rows, "x"]
    }

    U<-runif(N*R) 
    sim.data<-array(qkde(U,tbl),dim=c(N,R)) 
    
    
  }
  else if(mod=="Boot"){
    sim.data<-array(sample(data,N*R,replace = T),dim=c(N,R))
    
  }
  samp.dist<-apply(sim.data,2,sum.measure)
  return(samp.dist)
}

construct_ci<-function(samp.dist){
  return(quantile(samp.dist,c(.05,.95),na.rm=T))
}

capture_par<-function(ci,true.par){
  1*(ci[1]<true.par&true.par<ci[2])
}
```



# Run function and attain results

```{r}
N<-201
true.norm.med<-qnorm(.5)+4.1
true.norm.min<-mean(apply(array(rnorm(N*10000),dim=c(N,10000)),2,min))+4.1
true.gamma.med<-qgamma(.5,shape=1.4,scale=3)
true.gamma.min<-mean(apply(array(rgamma(N*10000,shape=1.4,scale=3),dim=c(N,10000)),2,min))


settings<-expand.grid(dist=c("norm","gamma"),model=c('MMnorm','MMgamma','KDE','Boot'),par.int=c('min','median'),stringsAsFactors = F)
settings<-settings %>% 
  mutate(true.par=case_when(
    dist=='norm'&par.int=='min'~true.norm.min,
    dist=='norm'&par.int=='median'~true.norm.med,
    dist=='gamma'&par.int=='min'~true.gamma.min,
    dist=='gamma'&par.int=='median'~true.gamma.med
  )) %>% 
  mutate(id=row_number()) %>% 
  select(id,everything())



cores_2_use <- detectCores() - 1
print('cores_2_use: ',cores_2_use)
cl <- makeCluster(cores_2_use)
clusterSetRNGStream(cl, 2344)
registerDoParallel(cl)

for (j in 1:nrow(settings)){
  captures <- foreach(
    i = 1:20
    , .combine = c
    , .packages = c('dplyr') 
  ) %dopar% {
    generate_data(N,settings$dist[j]) %>% estimate.ci(mod=settings$model[j],par.int=settings$par.int[j],R=5000 )%>%construct_ci() %>% capture_par(settings$true.par[j])
  }
  settings$cap[j]<-mean(captures)
}

stopCluster(cl)
```
```{r}
settings
```


# Conclusion
By calculating, we can know that not all methods can work. So we can see that some answers are not perfectly correct and their coverage probability is very low.

