---
title: "10-clt"
author: "Qianyi Sun"
output: 
  html_document:
    theme: united
    #highlight: zenburn
    code_folding: hide
    toc: true
    toc_depth: 3
    toc_float: true
    #number_sections: true
    fig_caption: true
---

# Introduction
he central limit theorem is an important computational short-cut for generating and making inference from the sampling distribution of the mean. You’ll recall that the central limit theorem short-cut relies on a number of conditions, specifically:

Independent observations
Identically distributed observations
Mean and variance exist
Sample size large enough for convergence

```{r}
library(magrittr)
library(sn)
```


```{r}
require(magrittr)
require(sn)


# Parameters that dont change
R <- 5000
location <- 0
scale <- 1

# Parameters that change
N <- 5
slant <- 10
delta <- slant/(sqrt(1+slant^2))

# Quantites to calculate/generate
pop_mean <- location+scale*delta*(sqrt(2/pi)) 
pop_sd <- sqrt(scale^2*(1-((2*delta^2)/pi)))

# Check if we create the distribution correctly
sample <- rsn(10000,xi=location,omega=scale,alpha=slant)
mean(sample)
pop_mean
sd(sample)
pop_sd

Z <- rnorm(R)

sample_dist_clt <- Z*(pop_sd/sqrt(N))+pop_mean
random.skew <- array(rsn(R*N,xi=location,omega=scale,alpha=slant),dim = c(R,N))
sample_dist_sim <- apply(random.skew, 1, mean)

# QQ plot
qqplot(sample_dist_sim,sample_dist_clt,asp = 1)
abline(0,1)

# Optional (Dislay the 95% CI from the clt and the sim)
```



